---
title: "WQD7004--Group Assignment--DataCleaning"
author: "Thoo Pooi Luen"
date: '2022-04-27'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Part 1. Data Cleaning

```{r}
## library 
library(dplyr)
library(tidyverse)
library(stringr)  # to trim dirty character 
```

## import data

### use sep = "," to seperate data into multiple column

```{r}
combined_data <- read.csv("/Users/Fanny/Documents/R/um-assignment/gp_code_repository/data/student-data.csv",
                          header = TRUE, 
                          sep = ",", 
                          quote = "", 
                          as.is = TRUE,
                          na.strings = "NULL",
                          stringsAsFactors = FALSE)

```

##  Select top 5 data list 
```{r}
head(combined_data,5)
```

## Print out data summary 
```{r}
summary(combined_data)
```

# determine data type of loaded data

```{r}
typeof(combined_data)
class(combined_data)
```

# convert as tibble, mini version of data frame
```{r}
combined_data_tibble <- as.tibble(combined_data)
head(combined_data_tibble,5)
```

# refine column name
## 1. print out list of column name
```{r}
names(combined_data_tibble)
```

## 2. Detect data type of column name
```{r}
typeof(names(combined_data_tibble))
```

## 3. Trim unwanted header value, replace them as table name, view updated table
```{r}
trimmed_col_name <- stri_trim_both(names(combined_data_tibble),
                                   pattern="[^(*.X.)$,]")
names(combined_data_tibble) <- trimmed_col_name
view(combined_data_tibble)
```

# refine row data format for each column
## 1. print out first 5 row of data
```{r}
head(combined_data_tibble,5)
```

## 2. trim unwanted value by row name
```{r}
# SOS weird output 
head(str_trim(combined_data_tibble),5)
```

## 3. prepare trim function to discard unwanted characters
```{r}
# SOS seeing error 
# i am handling column 12 - 22
list_item_number <- 12:22


n <- min(list_item_number)
while (n < max(list_item_number)){
  selected_row <- combined_data_tibble[[list_item_number+1]]
  output <- stri_trim_both(selected_row, pattern = '[^"\\$]')
  print(output)
  
  n <- n+1
  
  if (n == max(list_item_number)){
    selected_row <- combined_data_tibble[[list_item_number+1]]
    output <- stri_trim_both(selected_row, pattern = '[^"\\$"]')
    print(output)
    break
  }
}
```
