---
title: "Research On How To Use caret Package"
output: html_notebook
---

We would like to research on how to use `caret` to standardize our modeling process when working with different ML algorithms.

### References

Some learning resources, examples and troubleshooting guides:

- [The caret Package](https://topepo.github.io/caret/)
- [A Short Introduction to the caret Package](https://cran.r-project.org/web/packages/caret/vignettes/caret.html)
- [Caret Package – A Practical Guide to Machine Learning in R](https://www.machinelearningplus.com/machine-learning/caret-package/)
- [Machine Learning with caret in R](https://www.datacamp.com/courses/machine-learning-with-caret-in-r)
- [Error using library(caret) in RStudio](https://stackoverflow.com/questions/48416124/error-using-librarycaret-in-rstudio)
- [Predictive modelling of Iris Species with Caret](https://subhayo.wordpress.com/2019/06/12/predictive-modelling-with-caret-classification-of-iris-species/)

### Installation

Uncomment and run the installation code below in console. This will take a while.

```{r}
# install.packages("caret", dependencies = c("Depends", "Suggests"))
```

If encounter errors when loading `caret`, install these packages as well (or refer to the error messages). Otherwise, we can skip this step.

```{r}
# install.packages("gower")
# install.packages("hardhat")
# install.packages("parallelly")
```

Load the `caret` package.

```{r}
library(caret)
```

### Classification

We use `iris` dataset to illustrate a classification example.

#### Read Dataset

```{r}
df_iris <- iris
head(df_iris)
```

#### Check Structure

```{r}
str(df_iris)
```

Target column is already represented in factor.

#### Check Summary

```{r}
summary(df_iris)
```

We will skip data cleaning and EDA.

#### Perform Train-Test Split

We use `createDataPartition` function to create balanced splits.

```{r}
set.seed(42)

train_index <- createDataPartition(
  df_iris$Species,
  p = 0.8,
  list = FALSE,
  times = 1
)

train_set <- df_iris[train_index, ]
test_set <- df_iris[-train_index, ]
```

```{r}
nrow(train_set)
```

```{r}
nrow(test_set)
```

#### Modeling

We specify our resampling type using `trainControl` function - repeated 10-fold CV.

```{r}
fit_control <- trainControl(
  method = "repeatedcv",
  number = 10,
  repeats = 10
)
```

First, we fit our data with decision tree algorithm. We do not need hyperparameter tuning in this case. We use the `train` function to specify our arguments. Note that we also include the preprocessing step to standardize the data.

```{r}
set.seed(42)

dt_fit <- train(
  Species ~ .,
  data = train_set,
  method = "rpart",
  trControl = fit_control,
  preProcess = c("center", "scale")
)

dt_fit
```

We make predictions using our decision tree model.

```{r}
dt_pred <- predict(dt_fit, test_set)
```

We evaluate our decision tree model using the `confusionMatrix` function.

```{r}
confusionMatrix(dt_pred, test_set$Species)
```

Next, we try to fit a random forest model with hyperparameter tuning. We create a tuning parameter grid.

```{r}
rf_grid <- expand.grid(
  mtry = 2:4,
  splitrule = "gini",
  min.node.size = 1
)
```

We use the `train` function to specify our arguments. This time with our tuning parameter grid.

```{r}
set.seed(42)

rf_fit <- train(
  Species ~ .,
  data = train_set,
  method = "ranger",
  trControl = fit_control,
  preProcess = c("center", "scale"),
  tuneGrid = rf_grid
)

rf_fit
```

We make predictions using our random forest model.

```{r}
rf_pred <- predict(rf_fit, test_set)
```

We evaluate our random forest model using the `confusionMatrix` function.

```{r}
confusionMatrix(rf_pred, test_set$Species)
```

Alternatively, we can use `tuneLength` parameter.

```{r}
set.seed(42)

rf2_fit <- train(
  Species ~ .,
  data = train_set,
  method = "ranger",
  trControl = fit_control,
  preProcess = c("center", "scale"),
  tuneLength = 3
)

rf2_fit
```

We make predictions using our second random forest model.

```{r}
rf2_pred <- predict(rf2_fit, test_set)
```

We evaluate our second random forest model using the `confusionMatrix` function.

```{r}
confusionMatrix(rf2_pred, test_set$Species)
```
