---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

List of used library
```{r}
library(dplyr)
library(tidyverse)
library(ggplot2)
library(reshape2)
library(ggbeeswarm) 
```

# 1. Read processed dataset as data frame as 'student_por', and have a glimpse look at available attributes and factor

```{r}
student_por <- read.csv("C:/Users/Fanny/Documents/GitHub/programming-group-project/data/student-por_v3.csv", header = TRUE, sep = ",", stringsAsFactors = TRUE)
student_por <- data.frame(student_por)
str(student_por)
```

# 2. Mutate selected attributes' data type into factors that come with more instinctive representative value in dataset

### Remarks: Scale conversion reference -

### <https://davidfharris.com/2014/10/05/avoid-bias-poor-excellent-scale-making-balanced/>

```{r}
# Medu / Male Education in ordered factors manner
student_por$Medu  <- recode_factor(student_por$Medu,
                                   `0` = "none", 
                                   `1` = "primary_edu", 
                                   `2` = "5th_to_9th_grade", 
                                   `3` = "secondary_edu", 
                                   `4` = "higher_edu", .ordered = TRUE)

# Fedu / Female Education in ordered factors manner
student_por$Fedu  <- recode_factor(student_por$Fedu,
                                   `0` = "none", 
                                   `1` = "primary_edu", 
                                   `2` = "5th_to_9th_grade", 
                                   `3` = "secondary_edu", 
                                   `4` = "higher_edu", .ordered = TRUE)

# traveltime / home to school travel time 
student_por$traveltime  <- recode_factor(student_por$traveltime, 
                                         `1` = "<15 min", 
                                         `2` = "15 min - 30 min", 
                                         `3` = "30 min - 1 hour", 
                                         `4` = ">1 hour", .ordered = TRUE)

# studytime / weekly study time 
student_por$studytime  <- recode_factor(student_por$studytime, 
                                        `1` = "<2 hours", 
                                        `2` = "2 - 5 hours", 
                                        `3` = "5 - 10 hours", 
                                        `4` = ">10 hours", .ordered = TRUE)

# famrel / quality of family relationships
student_por$famrel  <- recode_factor(student_por$famrel, 
                                     `1` = "Very bad", 
                                     `2` = "Fair", 
                                     `3` = "Good", 
                                     `4` = "Very good", 
                                     `5` = "Excellent", .ordered = TRUE)

# freetime / free time after school
student_por$freetime  <- recode_factor(student_por$freetime, 
                                     `1` = "Very low", 
                                     `2` = "Low", 
                                     `3` = "Fair", 
                                     `4` = "High", 
                                     `5` = "Very high", .ordered = TRUE)

# goout / going out with friends
student_por$goout  <- recode_factor(student_por$goout, 
                                     `1` = "Very low", 
                                     `2` = "Low", 
                                     `3` = "Fair", 
                                     `4` = "High", 
                                     `5` = "Very high", .ordered = TRUE)

# Dalc  / workday alcohol consumption
student_por$Dalc  <- recode_factor(student_por$Dalc, 
                                     `1` = "Very low", 
                                     `2` = "Low", 
                                     `3` = "Fair", 
                                     `4` = "High", 
                                     `5` = "Very high", .ordered = TRUE)

# Walc  / weekend alcohol consumption
student_por$Walc  <- recode_factor(student_por$Walc, 
                                     `1` = "Very low", 
                                     `2` = "Low", 
                                     `3` = "Fair", 
                                     `4` = "High", 
                                     `5` = "very high", .ordered = TRUE)

# health / current health status
student_por$health  <- recode_factor(student_por$health, 
                                     `1` = "very bad", 
                                     `2` = "Somewhat bad", 
                                     `3` = "Fair", 
                                     `4` = "Somewhat good", 
                                     `5` = "very good", .ordered = TRUE)
```

# 3. Check on dataset summary

```{r}
summary(student_por)
```

# 4. Preparare General Function for Graph Plotting

```{r}
# bar graph for bivariate analysis
Unstacked_bi_bar_graph <- function(`x.axis` = "", Result = ""){
  # Result represents dependent variable / usually being drawn in y-axis 
  # x-axis represent independent variable 
  # dodge is used to un-stake the bar graph
  
  # 1. counts (or sums of weights)
  bar <- ggplot(student_por, aes(`x.axis`))
  
  # 2. Number of tuples in each class
  graph_output <- bar + geom_bar(aes(fill = `Result`), position = 'dodge') + 
    labs(y = 'Frequency') 
  
  # 3. get data from the graph 
  graph_label <- layer_data(graph_output)
  
  # 4. Annotate value at respective bar
  graph_output <- graph_output + annotate(geom = "text", label = graph_label$count, 
                                          x = graph_label$x, y = 15)
  
  return(graph_output)
}
```

# 5. Section 1: Bivariate Visualization for Expected Result Attribute

### a. School Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$school, student_por$result)
output = output + ggtitle(label = 'School versus Expected Result', 
                          subtitle = 'school vs result') +
  labs(x = "Student's School") 
output
```

### b. Sex Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$sex, student_por$result)
output = output + ggtitle(label = 'Gender versus Expected Result', 
                          subtitle = 'sex vs result') +
  labs(x = "Student's Sex") 
output
```

### c. Age Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$age, student_por$result)
output = output + ggtitle(label = 'Age versus Expected Result', 
                          subtitle = 'age vs result') +
  labs(x = "Student's Age") 
output

min(student_por$age)
max(student_por$age)
```

### d. Address Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$address, student_por$result)
output = output + ggtitle(label = "Student's Home Address Type versus Expected Result", 
                          subtitle = 'address vs result') +
  labs(x = "Address Type") 
output
```

### e. Family Size Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$famsize, student_por$result)
output = output + ggtitle(label = "Student's Family Size versus Expected Result", 
                          subtitle = 'famsize vs result') +
  labs(x = "Family Size") 
output
```

### f. Parent's Cohabitation Status Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$Pstatus, student_por$result)
output = output + ggtitle(label = "Parent's Cohabitation Status versus Expected Result", 
                          subtitle = 'Pstatus vs result') +
  labs(x = "Parent's Cohabitation Status") 
output
```

### g. Mother's Education Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$Medu, student_por$result)
output = output + ggtitle(label = "Mother's Education Vs Expected Result", 
                          subtitle = 'Medu vs result') +
  labs(x = "Mother's Education Level") 
output
```

### h. Father's Education Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$Fedu, student_por$result)
output = output + ggtitle(label = "Father's Education Vs Expected Result", 
                          subtitle = 'Fedu vs result') +
  labs(x = "Father's Education Level") 
output
```

### i. Mother's Job Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$Mjob, student_por$result)

output = output + ggtitle(label = "Mother's Job Vs Expected Result", 
                          subtitle = 'Mjob vs result') +
  labs(x = "Mother's Job") 
output
```

### j. Father's Job Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$Fjob, student_por$result)

output = output + ggtitle(label = "Father's Job Vs Expected Result", 
                          subtitle = 'Fjob vs result') +
  labs(x = "Father's Job") 
output
```

### k. Reason to Choose Selected School Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$reason, student_por$result)

output = output + ggtitle(label = "Reason to Choose Selected School Vs Expected Result", 
                          subtitle = 'reason vs result') +
  labs(x = "Reason Types") 
output
```

### l. Student's Guardian Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$guardian, student_por$result)

output = output + ggtitle(label = "Student's Guardian Vs Expected Result", 
                          subtitle = 'guardian vs result') +
  labs(x = "Guardian") 
output
```

### m. Home to School Travel Time Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$traveltime, student_por$result)

output = output + ggtitle(label = "Home to School Travel Time Vs Expected Result", 
                          subtitle = 'traveltime vs result') +
  labs(x = "Travel Time") 
output
```

### n. Weekly Study Time Vs Expected Result

```{r}
## the sample code i refer to is showing scatterplot, but when being implemented into dataset, become like this
# create scatter plot of hours studied vs. exam score
plot(student_por$studytime, student_por$result, pch=16, col='steelblue',
     main='Weekly Study Time Vs Expected Result',
     xlab='Weekly Study Time', ylab='Expected Result')
```

```{r}
output = Unstacked_bi_bar_graph(student_por$studytime, student_por$result)

output = output + ggtitle(label = "Weekly Study Time Vs Expected Result", 
                          subtitle = 'studytime vs result') +
  labs(x = "Weekly Study Time") 
output
```

### o. Number of Past Class Failures Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$failures, student_por$result)

output = output + ggtitle(label = "Number of Past Class Failures Vs Expected Result", 
                          subtitle = 'failures vs result') +
  labs(x = "Number of Past Class Failures") 
output
```

### p. Extra Educational Support Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$schoolsup, student_por$result)

output = output + ggtitle(label = "Extra Educational Support Vs Expected Result", 
                          subtitle = 'schoolsup vs result') +
  labs(x = "Extra Educational Support") 
output
```

### q. Family Educational Support Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$famsup, student_por$result)

output = output + ggtitle(label = "Family Educational Support Vs Expected Result", 
                          subtitle = 'famsup vs result') +
  labs(x = "Family Educational Support") 
output
```

### r. Extra Paid Classes Within The Course Subject(Portuguese) Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$paid, student_por$result)

output = output + ggtitle(label = "Extra Paid Classes for Portuguese Subject Vs Expected Result", 
                          subtitle = 'paid vs result') +
  labs(x = "Joined Extra Paid Classes?") 
output
```

### s. Extra-Curricular Activities Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$activities, student_por$result)

output = output + ggtitle(label = "Extra-Curricular Activities Vs Expected Result", 
                          subtitle = 'activities vs result') +
  labs(x = "Joined Extra-Curricular Activities?") 
output
```

### t. Intention to Pursue Higher Education Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$higher, student_por$result)

output = output + ggtitle(label = "Intention to Pursue Higher Education Vs Expected Result", 
                          subtitle = 'higher vs result') +
  labs(x = "Intention to Pursue Higher Education") 
output
```

### u. Internet Access at Home Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$internet, student_por$result)

output = output + ggtitle(label = "Internet Access at Home Vs Expected Result", 
                          subtitle = 'internet vs result') +
  labs(x = "Internet Access at Home") 
output
```

### v. With a Romantic Relationship Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$romantic, student_por$result)

output = output + ggtitle(label = "With a Romantic Relationship Vs Expected Result", 
                          subtitle = 'romantic vs result') +
  labs(x = "With a Romantic Relationship?") 
output
```

### w. Quality of Family Relationships Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$famrel, student_por$result)

output = output + ggtitle(label = "Quality of Family Relationships Vs Expected Result", 
                          subtitle = 'famrel vs result') +
  labs(x = "Quality of Family Relationships") 
output
```

### x. Free Time After School Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$freetime, student_por$result)

output = output + ggtitle(label = "Free Time After School Vs Expected Result", 
                          subtitle = 'freetime vs result') +
  labs(x = "Free Time After School") 
output
```

### y. Going Out with Friends Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$goout, student_por$result)

output = output + ggtitle(label = "Going Out with Friends Vs Expected Result", 
                          subtitle = 'goout vs result') +
  labs(x = "Going Out with Friends") 
output
```

### z. Workday Alcohol Consumption Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$Dalc, student_por$result)

output = output + ggtitle(label = "Workday Alcohol Consumption Vs Expected Result", 
                          subtitle = 'Dalc vs result') +
  labs(x = "Workday Alcohol Consumption") 
output
```

### aa. Weekend Alcohol Consumption Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$Walc, student_por$result)

output = output + ggtitle(label = "Weekend Alcohol Consumption Vs Expected Result", 
                          subtitle = 'Walc vs result') +
  labs(x = "Weekend Alcohol Consumption") 
output
```

### ab. Current Health Status Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$health, student_por$result)

output = output + ggtitle(label = "Current Health Status Vs Expected Result", 
                          subtitle = 'health vs result') +
  labs(x = "Current Health Status") 
output
```

### ac. Number of School Absences Vs Expected Result

```{r}
output = Unstacked_bi_bar_graph(student_por$absences, student_por$result)

output = output + ggtitle(label = "Number of School Absences Vs Expected Result", 
                          subtitle = 'absences vs result') +
  labs(x = "Number of School Absences") 
output
```

```{r}
# beeswarm 
ggplot(student_por, aes(x = result, y = absences, color = result)) +
  geom_beeswarm(cex = 3, priority = "density") + 
  ggtitle(label = "Number of School Absences Vs Expected Result", 
          subtitle = 'absences vs result') +
  labs(x = "Number of School Absences", y = "Marks") 
```

### ad. First Period Grade Vs Expected Result

```{r}
# boxplot creation
ggplot(student_por, aes(x = G1, y = result)) +
  geom_boxplot() + 
  labs(title = "First Period Grade Vs Expected Result",
       subtitle = 'G1 vs result',
       x = "First Period Grade", 
       y = 'Result')
```

```{r}
# this is how the graphs looks like without separation by result 
output = Unstacked_bi_bar_graph(student_por$G1, student_por$result)

output = output + ggtitle(label = "First Period Grade Vs Expected Result", 
                          subtitle = 'G1 vs result') +
  labs(x = "First Period Grade") 
output
```

```{r}
# beeswarm 
ggplot(student_por, aes(x = result, y = G1, color = result)) +
  geom_beeswarm(cex = 3, priority = "density") + 
  ggtitle(label = "First Period Grade Vs Expected Result", 
          subtitle = 'G1 vs result') +
  labs(x = "First Period Grade", y = "Marks") 
```

### ae. Second Period Grade Vs Expected Result

```{r}
# boxplot creation
ggplot(student_por, aes(x = G2, y = result)) +
  geom_boxplot() + 
  labs(title = "Second Period Grade Vs Expected Result",
       subtitle = 'G2 vs result',
       x = "Second Period Grade", 
       y = 'Result')
```

```{r}
# beeswarm 
ggplot(student_por, aes(x = result, y = G2, color = result)) +
  geom_beeswarm(cex = 3, priority = "density") + 
  ggtitle(label = "Second Period Grade Vs Expected Result", 
          subtitle = 'G2 vs result') +
  labs(x = "Second Period Grade", y = "Marks") 
```

```{r}
# counts (or sums of weights)
bivariate_bargraph <- ggplot(student_por, aes(G2))

# Number of tuples in each class:
bivariate_bargraph + geom_bar(aes(fill = result), position = 'dodge') + 
  #facet_wrap(~ result) + 
  labs(title = "Second Period Grade Vs Expected Result",
       subtitle = 'G2 vs result',
       x = "Second Period Grade", 
       y = 'count')
```

### af. Final Grade Vs Expected Result

```{r}
# beeswarm 
ggplot(student_por, aes(x = result, y = G3, color = result)) +
  geom_beeswarm(cex = 3, priority = "density") + 
  ggtitle(label = "Final Grade Vs Expected Result", 
          subtitle = 'G3 vs result') +
  labs(x = "Final Grade", y = "Marks") 
```

# 6. Correlation Matrix

```{r}
# i not sure will we make R bias by converting attribute back to numeric form for chi-square test 
print("Pearson Correlation Coefficient between School and Result")
cor(as.numeric(student_por$school), as.numeric(student_por$result))

print("Correlation Between between School and Result")
cor.test(as.numeric(student_por$school), as.numeric(student_por$result), method = c("pearson"), conf.level = 0.95)
```

# Random plot
```{r}
# Bivariate density plots
# In a 2D density plot, nested contours (or contours plus colors) indicate regions of higher local density. 
ggplot(student_por, aes(x = G3, y = absences)) + 
  geom_density2d() + 
  geom_jitter(alpha=0.35) + # using geom_jitter to avoid overplotting of points
  labs(title = "Final Grade vs Absences",
       subtitle = 'G3 vs absences',
       x = "Final Grade", 
       y = 'Absences') +
  theme_classic()  + 
  theme(aspect.ratio = 1)
```